\addchap{About the Roosevelt Club}

   \initial{O}ur world is facing pressing issues, to which orthodox approaches have
   yielded only unfruitful answers - we must surpass this shortcoming. Our
   work brings together individuals from differing backgrounds, uniting
   them in synergistic conversation and written reflection; our aim is to
   produce bold, innovative, and pragmatic thinking on the challenges of
   the modern era, and to prepare our Members for a life in service of the
   public good.

   Never before has such singular progress been made toward a better
   world. We live in a world of constant improvement -- to industry, to
   technology, to all the mechanisms that drive our society ever forward.
   The knowledge that we have amassed over the course of millennia is now
   readily available to anyone with access to an internet connection; and
   this knowledge is growing exponentially, stretching well beyond the
   limits of human understanding. We have reached heights of expertise and
   capability completely unimagineable mere decades in the past; we are
   able to grow back limbs, cure diseases that years ago would have
   decimated us; we have set foot on the Moon; observed and recorded
   hundreds of millions of galaxies, each containing celestial objects
   billions of times larger than our Earth.

   And yet never before has there been such radical and widespread
   inequality; never before have we been closer to scarcity. The bees that
   pollinate our crops are dying, our seemingly endless reserves of
   natural spring water are being depleted, our livestock is diseased and
   kept alive almost miraculously by a cocktail of medicines. Never before
   have we been closer to man-made environmental catastrophe; never before
   have we been threatened by such destructive weaponry.

   This is the world we live in: a world of constant dichotomy, constant
   uncertainty and constant peril. This is why we are at a unique moment
   in history. Now, and only now, do we exist in this balance: we have
   achieved so much, just enough to become aware that there remains so
   much to be achieved. And we must achieve it.

   We must overcome political squabbles in the face of the issues that are
   severe enough to break us. We must eliminate poverty and homelessness.
   We must push for equal rights amongst genders; and we must extend human
   rights to those in lack of them. We must protect our environment while
   developing our industry to the needs of a new economy, and we must
   eliminate the dangers to our food production. We must find renewable
   alternatives to our natural resources while increasing their
   availability. We must not extend aid, but eliminate the need for it.

   This task is undoubtedly a daunting one, but it comes down to us. This
   is the task of our generation: to overcome the failures of our
   predecessors, and to secure our progress toward a better world. It is
   larger than life, but it can be accomplished: the power that each of us
   holds is inconceivable, and its underestimation is what limits us; but
   it is there, latent, ready to be exercised. Now is when we hold this
   power, and now is when we must exercise it.

   This is what we believe in, and this is what we are working toward. We
   want to gather those who are passionate about these issues, passionate
   about enacting change, and push the limits of what these individuals
   can achieve. We want to voice our concerns about the picture we paint
   of the world and offer, in its place, alternative, solution-oriented
   ideas on how we can tackle the most pressing issues of our time.


