\addchap{Letter From The Editor}


\initial{T}he guiding ethos of our journal is to reach for a broader, longer-view perspective on the political events that define the world around us. This November marks the 30th anniversary of the fall of the Berlin Wall---by any account, one of the most consequential events of the 20th century. In retrospect, the conflict between West and 'Evil Empire' may not seem as existential; at the very least, we know that the West was bound to win: the Soviet Union's economy had been in steady decline for nearly three decades, it did not posses the physical or human capital required to catch up, and was plagued by its internal politics.  

At the time, the picture was quite different: the Soviet Union was a monolithic titan, and the fall of the Berlin Wall represented the culmination of a formidable effort---and was a momentous victory indeed. Then came the post-war honeymoon (I won't do you the disservice of quoting Fukuyama, again). This, of course, was not all that we would have hoped, and the promise of an open world has been rudely abandoned. 

As we reach this thirty year mark, our world once again inches towards harder borders, taller walls, and deeper divisions. In this much-awaited second issue, our Members reflect on these past thirty years, and on what walls have meant throughout history. 

While we gear up for another pivotal year in politics, we hope this issue will give you the necessary to pause to think, perhaps just a little more deeply, about where our world is going. \\

Sincerely,

Laurent B\'elanger-Lowe